---
title: "Course Project 1 - Reproducible Research"
author: "daniela di michele"
date: "dicembre 2019"
output:
  html_document: default
  pdf_document: default
---

Step 1 - Code for reading in the dataset and/or processing the data
```{r}
library(lubridate)
library(dplyr)
activity<-read.csv("C:/Users/daniela.dimichele/Desktop/coursera/corso_5/test_2/activity.csv",header=TRUE,sep=",")
summary(activity)
activity$date<-ymd(activity$date)
```
INFORMATION ABOUT THE VARIABLES
```{r}
dim(activity)
head(activity)
str(activity)
groupActivity <- group_by(activity, date)
Activity_2<-summarize(groupActivity, steps= sum(steps, na.rm = TRUE ))
Activity_2
```
##Step 2 - Histogram of the total number of steps taken each day
```{r}
hist(Activity_2$steps,
    main = "Histogram of the Total Number of Steps Taken per Day", 
    xlab = "Number of Steps", 
    ylab = "Number of days",
    col  = "Green",
    breaks=10)
abline(v = mean(Activity_2$steps), lty = 1, lwd = 2, col = "orange")
abline(v = median(Activity_2$steps), lty = 2, lwd = 2, col = "black")
legend(x = "topleft", c("Mean", "Median"), col = c("orange", "black"), 
       lty = c(1, 2), lwd = c(2, 2))

```
##Step 3 - Mean and median number of steps taken each day
```{r}
mean(Activity_2$steps)
median(Activity_2$steps)
```

##Step 4 - Time series plot of the average number of steps taken
```{r}
Group_Activity <- group_by(activity, interval)
Activity_mean<-summarize(Group_Activity, steps= mean(steps, na.rm = TRUE ))
Activity_mean

plot(Activity_mean$interval, Activity_mean$steps, ty = "l",
     col="green",
     xlab = "Time interval", 
     ylab = "Mean Steps Every day", 
     main = "Mean Steps by Date")
```

##Step 5 - The 5-minute interval that, on average, contains the maximum number of steps
```{r}
Activity_mean[which.max(Activity_mean$steps), ]$interval
summary(activity)
```
##Step 6 - Code to describe and show a strategy for imputing missing data
```{r}
activity_new<-activity
for (i in 1:nrow(activity_new))
{
  if (is.na(activity_new$steps[i])){
    activity_new$steps[i]<-Activity_mean[which(activity_new$interval[i] == Activity_mean$interval),]$steps
  }
}
summary(activity_new)
sum(is.na(activity_new))
group_activity_new <- group_by(activity_new, date)
data_fi<-summarize(group_activity_new, steps= sum(steps, na.rm = TRUE ))
data_fi
```

##Step 7 - Histogram of the total number of steps taken each day after missing values are imputed
```{r}
hist(data_fi$steps, breaks=10,
     col = "red",
     main = "Total number of steps x day", 
     xlab = "Steps",ylab="Frequency")
mean(data_fi$steps)
median(data_fi$steps)
``` 

```{r}
par(mfrow=c(1,2))
hist(activity_new$steps, 
     breaks=10,
     main = "Total number of steps per day", 
     xlab = "Steps", 
     col = "red")
hist(data_fi$steps, 
     breaks=10,
     main = "Total number of steps x day with no NA's", 
     xlab = "Steps",
     col = "red")
```

##Step 8 - Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r}
library(ggplot2)

for (i in 1:nrow(activity))
{
  if(weekdays(activity$date[i])=="Saturday"|weekdays(activity$date[i])=="Sunday")
    { 
    activity$weekdays[i]<-"weekend"
  }else{activity$weekdays[i]<-"weekday"}
}
inte_group_Na <- group_by(activity, interval,weekdays)
dataNA<-summarize(inte_group_Na, steps= mean(steps, na.rm = TRUE ))
qplot(interval,steps, data = dataNA,facets = weekdays~., geom="line")
```


